Highly Available Transactions: Virtues and Limitations\iftechreport\\(Extended Version)\fi

### 2013-10-02 writeup/conclusion.tex
More editing

The current state of database software offers uncomfortable and
unnecessary choices between availability and transactional semantics.
Through our analysis and experiments, we have demonstrated how goals
of high availability will remain a critical aspect of many future data
storage systems. Accordingly, we expose a broad design space of Highly
Available Transactions (HATs), which can offer the key benefits of
highly available distributed systems---``always on'' operation during
partitions and low-latency operations---while also providing a family
of transactional isolation levels and replica semantics that have been
adopted in practice.  In taxonomizing this design space, we show that
many transactional guarantees are achievable with high availability,
while a few, like preventing Lost Update and Write Skew, are
not. Surprisingly, many of the default (and sometimes strongest)
semantics provided by today's traditional database systems are
achievable as HATs, hinting that distributed databases need not
compromise availability in order to serve many existing applications.

HATs raise several challenges for future study. In this paper, we have
largely focused on previously defined isolation and data consistency
models from the consistency from the database and distributed systems
communities. Their previous definitions and, in many cases, widespread
adoption hints at their utility to end-users. However, we believe
there is considerable work to be done to improve the programmability
of highly available systems. Isolation and data consistency are means
by which \textit{application-level} consistency is achieved but are
typically not end goals for end-user applications. Our results hint
that a mix of HAT and non-HAT semantics (the latter used sparingly) is
required for practical applications, but the decision to employ each
and the system architectures for a hybrid approach remain open
problems. While we have studied the analytical and experiment behavior
of several HAT models, there is substantial work in further
understanding the performance and design of systems within the large
set of HAT models. Weakened failure assumptions as in escrow or in the
form of bounded network asynchrony could enable richer HAT semantics
at the cost of general-purpose availability. Alternatively, there is a
range of possible solutions providing strong semantics during
partition-free periods and weakened semantics during partitions. Based
on our understanding of what is desired in the field and newfound
knowledge of what is possible to achieve, we believe HATs provide a
large and useful design space for exploration.

>>>>

The current state of database software offers uncomfortable and
unnecessary choices between availability and transactional semantics.
Through our analysis and experiments, we have demonstrated how goals
of high availability will remain a critical aspect of many future data
storage systems. We expose a broad design space of Highly Available
Transactions (HATs), which can offer the key benefits of highly
available distributed systems---``always on'' operation during
partitions and low-latency operations (often orders of magnitude lower
than non-compliant)---while also providing a family of transactional
isolation levels and replica semantics that have been adopted in
practice.  We also identify many semantic guarantees that are
unachievable with high availability, including Lost Update and Write
Skew anomaly prevention, concurrent update prevention, and bounds on
data recency. Despite these limitations, and somewhat surprisingly,
many of the default (and sometimes strongest) semantics provided by
today's traditional database systems are achievable as HATs, hinting
that distributed databases need not compromise availability, low
latency, or scalability in order to serve many existing applications.

In this paper, we have largely focused on previously defined isolation
and data consistency models from the consistency from the database and
distributed systems communities. Their previous definitions and, in
many cases, widespread adoption hints at their utility to
end-users. However, we believe there is considerable work to be done
to improve the programmability of highly available systems. Isolation
and data consistency are means by which \textit{application-level}
consistency is achieved but are typically not end goals for end-user
applications. Our results hint that a mix of HAT and non-HAT semantics
(the latter used sparingly) is required for practical applications,
but the decision to employ each and the system architectures for a
hybrid approach remain open problems. While we have studied the
analytical and experiment behavior of several HAT models, there is
substantial work in further understanding the performance and design
of systems within the large set of HAT models. Weakened failure
assumptions as in escrow or in the form of bounded network asynchrony
could enable richer HAT semantics at the cost of general-purpose
availability. Alternatively, there is a range of possible solutions
providing strong semantics during partition-free periods and weakened
semantics during partitions. Based on our understanding of what is
desired in the field and newfound knowledge of what is possible to
achieve, we believe HATs represent a large and useful design space for
exploration.

### 2013-10-02 writeup/evaluation.tex
More editing

With an understanding of which semantics are HAT-compliant, in this
section, we analyze the implications of these results for existing
systems and briefly study HAT systems on public cloud
infrastructure. Specifically:

\begin{myenumerate}\vspace{-.5em}
\item We revisit traditional database concurrency control with a focus
  on coordination costs and on high availability.
\item We examine the properties required by an OLTP application based
  on the TPC-C benchmark.
\item We perform a brief experimental evaluation of HAT versus non-HAT
  properties on public cloud infrastructure.
\end{myenumerate}

\subsection{Existing Algorithms}
\label{sec:eval-existing}

Most existing database transaction and concurrency control algorithms
are not designed for high availability and often presume a
single-server deployment or a requirement for serializability. In this
section, we briefly discuss design decisions and algorithmic details
that preclude high availability.

@@

\vspace{.5em}\noindent\textbf{Non-serializability} Most existing
distributed implementations of weak isolation are not highly
available. Lock-based mechanisms such as those in Gray's original
proposal~\cite{gray-isolation} do not degrade gracefully in the
presence of partial failures. (Note, however, that lock-based
protocols \textit{do} offer the benefit of recency guarantees.) While
multi-versioned storage systems allow for a variety of transactional
guarantees, few offer traditional weak isolation (e.g.,
non-``tentative update'' schemes) in this context.  Chan and Gray's
read-only transactions have item-cut isolation with causal consistency
and MAV (session \textit{PL-2L}~\cite{adya}) but are unavailable in the
presence of coordinator failure and assume serializable update
transactions~\cite{readonly}; this is similar to read-only and
write-only transactions more recently proposed by
Eiger~\cite{eiger}.  Brantner's S3 database~\cite{kraska-s3}
and Bayou~\cite{sessionguarantees} can all provide variants of session
\textit{PL-2L} with high availability, but none provide this HAT
functionality without substantial modification. Swift~\cite{swift} and
bolt-on causal consistency~\cite{bolton} are closest to providing
maximal sticky HAT semantics. Accordingly, it is possible to implement
many guarantees weaker than serializability---including guarantees
that are achievable with high availability---and still not achieve
high availability.

@@

Thus far, we have largely ignored the question of when HAT semantics
are useful (or otherwise are too weak). As we showed in
Section~\ref{sec:hats}, the main cost of high availability and low
latency comes in the inability to prevent Lost Update, Write Skew, and
provide recency bounds. In this section, we attempt to understand when
these guarantees matter in a representative transactional application
based on the TPC-C benchmark~\cite{tpcc}.

\vspace{.5em}\noindent\textbf{TPC-C} To better understand the impact
of HAT-compliance in an application context, we consider a concrete
application: the TPC-C benchmark. In brief, we find that four of five
transactions can be executed with HATs, while the fifth may require
unavailability.

@@

To further understand the performance implications of HAT guarantees
in a real-world environment, we implemented a HAT database
prototype. Across a range of deployments on public cloud
infrastructures, we found that, as Section~\ref{sec:latency}'s
measurements suggested, ``strongly consistent'' algorithms incur
substantial latency penalties: over WAN, 10 to 100 times higher than
their HAT counterparts. We find that HAT
properties---while more resource-intensive than basic eventual
consistency---scale well to many servers.

\vspace{.5em}\noindent\textit{Implementation.} Our prototype database
is a partially replicated (hash-based, partitioned) key-value backed
by LevelDB and implemented in Java using Apache Thrift. It currently
supports eventual consistency (hereafter, \texttt{eventual};
last-writer-wins RU with standard all-to-all anti-entropy between
replicas) and the efficient HAT MAV algorithm as sketched in
Section~\ref{sec:ta}. (hereafter, \texttt{MAV}). It also supports
non-HAT operation whereby all operations for a given key are routed to
a (randomly) designated \texttt{master} replica for each key
(guaranteeing single-key linearizability, as in Gilbert and Lynch's
CAP Theorem proof~\cite{gilbert-cap} and in PNUTS~\cite{pnuts}'s
\texttt{read-latest} operation; hereafter, \texttt{master}) as well as
distributed two-phase locking. Servers are durable: they synchronously
write to LevelDB before responding to client requests, while new
writes in MAV are sent to a write-ahead log, which is synchronously
flushed to disk.

\vspace{.5em}\noindent\textit{Configuration.} We deploy the database
in \textit{clusters}---disjoint sets of database servers that each
contain a single, fully replicated copy of the data---typically across
datacenters, sticking all clients within a datacenter to their
respective cluster (trivially providing read-your-writes and monotonic
reads guarantees). By default, we use 5 Amazon EC2 \texttt{m1.xlarge}
instances as servers in each cluster. For our workload, we link our
client library to the YCSB benchmark~\cite{ycsb}, which is well-suited
to LevelDB's key-value schema, grouping every eight YCSB operations
from the default workload (50\% reads, 50\% writes) to form a
transaction. We increase the number of keys in the workload from the
default 1,000 to 100,000 with uniform random key access, keeping
the default value size of $1KB$, and running YCSB for 180 seconds per
configuration.

@@

\vspace{.5em}\noindent\textit{Transaction length.} As shown in
Figure~\ref{fig:txlen} (clusters in Virginia and Oregon), throughput
of \texttt{eventual}, \texttt{RC}, and \texttt{master} operation are
unaffected by transaction length. In contrast, \texttt{MAV} throughput
decreases linearly with increased transaction length: with $1$
operation per transaction, \texttt{MAV} throughput is within 18\% of
\texttt{eventual} ($34$ bytes overhead), and with $128$ operations per
transaction, \texttt{MAV} throughput is within $60\%$ ($1898$ bytes
overhead). This reflects our \texttt{MAV} algorithm's metadata requirements,
which are proportional to transaction length and consume IOPS and
network bandwidth.

@@

\vspace{.5em}\noindent\textit{Scale-out.} One of the key benefits of
our HAT algorithms is that they are shared-nothing, meaning they
should not compromise scalability. Figure~\ref{fig:scaleout} shows
that varying the number of servers across two clusters in Virginia and
Oregon (with $15$ YCSB clients per server) results in linear scale-out
for \texttt{eventual}, \texttt{RC}, and \texttt{MAV}. \texttt{RC} and
\texttt{eventual} scale perfectly with an approximately $5$x
throughput increase moving from $5$ to $25$ servers per cluster. For
the same configuration, \texttt{MAV} scales by $3.8$x, achieving over
$260,000$ operations per second. \texttt{MAV} performance is due to
resource contention---with a memory-backed database (instead of
LevelDB), \texttt{MAV} scales by $4.25$x (not shown)---and MAV-related
performance heterogeneity across servers (Calvin's authors report
similar heterogeneity on EC2~\cite{calvin}). Nevertheless, this linear
scalability is not typical of traditional database system
designs~\cite{gray-isolation}.

>>>>

With an understanding of which semantics are HAT-compliant, in this
section, we analyze the implications of these results for existing
systems and briefly study HAT systems on public cloud
infrastructure. Specifically:
\begin{myenumerate}
\item We revisit traditional database concurrency control with a focus
  on coordination costs and on high availability.
\item We examine the properties required by an OLTP application based
  on the TPC-C benchmark.
\item We perform a brief experimental evaluation of HAT versus non-HAT
  properties on public cloud infrastructure.
\end{myenumerate}

\subsection{HA and Existing Algorithms}
\label{sec:eval-existing}

While we have shown that many database isolation levels are achievable
as HATs, many traditional concurrency control mechanisms do not
provide high availability---even for HAT-compliant isolation
levels. Existing mechanisms often presume (or are adapted from)
single-server non-partitioned deployments or otherwise focus on
serializability as a primary use case. In this section, we briefly
discuss design decisions and algorithmic details that preclude high
availability.

@@

\vspace{.5em}\noindent\textbf{Non-serializability} Most existing
distributed implementations of weak isolation are not highly
available. Lock-based mechanisms such as those in Gray's original
proposal~\cite{gray-isolation} do not degrade gracefully in the
presence of partial failures. (Note, however, that lock-based
protocols \textit{do} offer the benefit of recency guarantees.) While
multi-versioned storage systems allow for a variety of transactional
guarantees, few offer traditional weak isolation (e.g.,
non-``tentative update'' schemes) in this context.  Chan and Gray's
read-only transactions have item-cut isolation with causal consistency
and MAV (session \textit{PL-2L}~\cite{adya}) but are unavailable in
the presence of coordinator failure and assume serializable update
transactions~\cite{readonly}; this is similar to read-only and
write-only transactions more recently proposed by Eiger~\cite{eiger}.
Brantner's S3 database~\cite{kraska-s3} and
Bayou~\cite{sessionguarantees} can all provide variants of session
\textit{PL-2L} with high availability, but none provide this HAT
functionality without substantial modification. Swift~\cite{swift} and
bolt-on causal consistency~\cite{bolton} are closest to providing the
maximal sticky HAT semantics we have considered. Accordingly, it is
possible to implement many guarantees weaker than
serializability---including guarantees that are achievable with high
availability---and still not achieve high availability. We view high
availability as a core design consideration in future concurrency
control designs.

@@

Thus far, we have largely ignored the question of when HAT semantics
are useful (or otherwise are too weak). As we showed in
Section~\ref{sec:hats}, the main cost of high availability and low
latency comes in the inability to prevent Lost Update, Write Skew, and
provide recency bounds. To better understand the impact of
HAT-compliance in an application context, we consider a concrete
application: the TPC-C benchmark. In brief, we find that four of five
transactions can be executed via HATs, while the fifth requires
unavailability.

@@

To further understand the performance implications of HAT guarantees
in a real-world environment, we implemented a HAT database
prototype. Here, we verify that, as Section~\ref{sec:latency}'s
measurements suggested, ``strongly consistent'' algorithms incur
substantial latency penalties (over WAN, 10 to 100 times higher than
their HAT counterparts) compared to HAT-compliant algorithms, which
scale linearly.
x
\vspace{.5em}\noindent\textit{Implementation.} Our prototype database
is a partially replicated (hash-based partitioned) key-value backed by
LevelDB and implemented in Java using Apache Thrift. It currently
supports eventual consistency (hereafter, \texttt{eventual};
last-writer-wins RU with standard all-to-all anti-entropy between
replicas) and the efficient HAT MAV algorithm as sketched in
Section~\ref{sec:ta}. (hereafter, \texttt{MAV}). We support non-HAT
operation whereby all operations for a given key are routed to a
(randomly) designated \texttt{master} replica for each key
(guaranteeing single-key linearizability, as in Gilbert and Lynch's
CAP Theorem proof~\cite{gilbert-cap} and in PNUTS~\cite{pnuts}'s
``read latest'' operation; hereafter, \texttt{master}) as well as
distributed two-phase locking. Servers are durable: they synchronously
write to LevelDB before responding to client requests, while new
writes in MAV are synchronously flushed to a disk-resident write-ahead
log.

\vspace{.5em}\noindent\textit{Configuration.} We deploy the database
in \textit{clusters}---disjoint sets of database servers that each
contain a single, fully replicated copy of the data---typically across
datacenters and stick all clients within a datacenter to their
respective cluster (trivially providing read-your-writes and monotonic
reads guarantees). By default, we deploy 5 Amazon EC2
\texttt{m1.xlarge} instances as servers in each cluster. For our
workload, we link our client library to the YCSB
benchmark~\cite{ycsb}, which is well-suited to LevelDB's key-value
schema, grouping every eight YCSB operations from the default workload
(50\% reads, 50\% writes) to form a transaction. We increase the
number of keys in the workload from the default 1,000 to 100,000 with
uniform random key access, keeping the default value size of $1KB$,
and running YCSB for 180 seconds per configuration.

@@

\vspace{.5em}\noindent\textit{Transaction length.} As shown in
Figure~\ref{fig:txlen} (clusters in Virginia and Oregon), throughput
of \texttt{eventual}, \texttt{RC}, and \texttt{master} operation are
unaffected by transaction length. In contrast, \texttt{MAV} throughput
decreases linearly with increased transaction length: with $1$
operation per transaction, \texttt{MAV} throughput is within 18\% of
\texttt{eventual} ($34$ bytes overhead), and with $128$ operations per
transaction, \texttt{MAV} throughput is within $60\%$ ($1898$ bytes
overhead). This reflects our \texttt{MAV} algorithm's metadata
requirements, which are proportional to transaction length and consume
IOPS and network bandwidth. We are currently investigating alternative
HAT algorithms that do not incur this overhead.

@@

\vspace{.5em}\noindent\textit{Scale-out.} One of the key benefits of
our HAT algorithms is that they are shared-nothing, meaning they
should not compromise scalability. Figure~\ref{fig:scaleout} shows
that varying the number of servers across two clusters in Virginia and
Oregon (with $15$ YCSB clients per server) results in linear scale-out
for \texttt{eventual}, \texttt{RC}, and \texttt{MAV}. \texttt{RC} and
\texttt{eventual} scale perfectly with an approximately $5$x
throughput increase moving from $5$ to $25$ servers per cluster. For
the same configuration, \texttt{MAV} scales by $3.8$x, achieving over
$260,000$ operations per second. \texttt{MAV} performance is due to
resource contention---with a memory-backed database (instead of
LevelDB), \texttt{MAV} scales by $4.25$x (not shown)---and MAV-related
performance heterogeneity across servers (Calvin's authors report
similar heterogeneity on EC2~\cite{calvin}). Our early experiments
with a newer database prototype with more efficient (non-Thrift) RPC
and connection pooling suggest that this scalability can be
substantially reduced. Nevertheless, this linear scalability is not
typical of traditional database system designs~\cite{gray-isolation}.

### 2013-10-02 writeup/hat-analysis.tex
More editing

>>>>

In light the of current practice of deploying weak isolation levels
(Section~\ref{sec:modernacid}), it is perhaps surprising that so many
weak isolation levels are achievable as HATs. Indeed, isolation levels
such as Read Committed expose and are defined in terms of end-user
anomalies that could not arise during serializable execution. However,
the prevalence of these models suggests that, in many cases,
applications can tolerate these their associated anomalies. Given our
HAT-compliance results, this in turn hints that--despite
idiosyncrasies relating to concurrent updates and data recency--highly
available database systems can provide sufficiently strong semantics
for many applications. Indeed, HAT databases may expose more anomalies
than a single-site database operating under weak isolation
(particularly during network partitions). However, for a fixed
isolation level (which, in practice, can vary across databases and may
differ from implementation-agnostic definitions in the literature),
users of single-site database are subject to the same (worst-case)
application-level anomalies as a HAT implementation. The necessary
(indefinite) visibility penalties (i.e., the right side of
Figure~\ref{fig:hat-order}) and lack of support for preventing
concurrent updates (via the upper left half of
Figure~\ref{fig:hat-order}) mean HATs are \textit{not} well-suited for
all applications (see Section~\ref{sec:evaluation}): these limitations
are fundamental. However, common practices such as ad-hoc, user-level
compensation and per-statement isolation ``upgrades'' (e.g.,
\texttt{SELECT FOR UPDATE} under weak isolation)--commonly used to
augment weak isolation--are also applicable in HAT systems (although
they may in turn compromise availability).

### 2013-10-02 writeup/hat-vldb.tex
More editing

   \vspace{0mm}
   \newcounter{qcounter}
    \begin{list}{\arabic{qcounter}.~}{\usecounter{qcounter}\leftmargin=1em}
        \setlength{\topsep}{0em}
        \setlength{\parskip}{0pt}
        \setlength{\partopsep}{0pt}
        \setlength{\parsep}{0pt}         
        \setlength{\itemsep}{.25em} 
        \setlength{\itemindent}{0em}
}
{
    \end{list} 
    \vspace{-.5em}
}

@@

\begin{abstract}
\noindent To minimize network latency and remain online during server
failures and network partitions, many modern distributed data storage
systems eschew transactional functionality, which provides strong
semantic guarantees for groups of multiple operations over multiple
data items. In this work, we consider the problem of providing Highly
Available Transactions (HATs): transactional guarantees that do not
suffer unavailability during system partitions or incur high network
latency.  We introduce a new taxonomy of highly available systems and
analyze existing ACID and distributed consistency guarantees to
identify which can and cannot be achieved in HAT systems. This unifies
the literature on weak transactional isolation, replica consistency,
and highly available systems. We analytically and experimentally
quantify the availability and performance benefits of HATs---often two
to three orders of magnitude over wide-area networks---as well as
their necessary semantic compromises.
\end{abstract}

@@

>>>>

   \vspace{-.5em}
   \newcounter{qcounter}
    \begin{list}{\arabic{qcounter}.~}{\usecounter{qcounter}\leftmargin=1em}
        \setlength{\topsep}{0em}
        \setlength{\parskip}{0pt}
        \setlength{\partopsep}{0pt}
        \setlength{\parsep}{0pt}         
        \setlength{\itemsep}{.25em} 
        \setlength{\itemindent}{0em}
}
{
    \end{list} 
    \vspace{-.5em}
}

@@

\begin{abstract}
\noindent To minimize network latency and remain online during server
failures and network partitions, many modern distributed data storage
systems eschew transactional functionality, which provides strong
semantic guarantees for groups of multiple operations over multiple
data items. In this work, we consider the problem of providing Highly
Available Transactions (HATs): transactional guarantees that do not
suffer unavailability during system partitions or incur high network
latency.  We introduce a taxonomy of highly available systems and
analyze existing ACID isolation and distributed data consistency
guarantees to identify which can and cannot be achieved in HAT
systems. This unifies the literature on weak transactional isolation,
replica consistency, and highly available systems. We analytically and
experimentally quantify the availability and performance benefits of
HATs---often two to three orders of magnitude over wide-area
networks---and discuss their necessary semantic compromises.
\end{abstract}

@@

\section*{Acknowledgments}
We would like to thank Peter Alvaro, Neil Conway, Evan Jones, Adam
Oliner, Aurojit Panda, and Shivaram Venkataraman for their helpful
feedback on this work. This research was supported in part by the Air
Force Office of Scientiﬁc Research (grant FA95500810352), DARPA XData
Award FA8750-12-2-0331, the Natural Sciences and Engineering Research
Council of Canada, the National Science Foundation (grants
CNS-0722077, IIS-0713661, and IIS-0803690), NSF CISE Expeditions award
CCF-1139158, the National Science Foundation Graduate Research
Fellowship (grant DGE-1106400), and gifts from Amazon Web Services,
Google, SAP, Cisco, Clearstory Data, Cloudera, Ericsson, Facebook,
FitWave, General Electric, Hortonworks, Huawei, Intel, Microsoft,
NetApp, Oracle, Samsung, Splunk, VMware, WANdisco, and Yahoo!.

### 2013-10-02 writeup/intro.tex
More editing

The last decade has seen a shift in the design of popular large-scale
database systems, from the use of transactional
RDBMSs~\cite{bernstein-book, gray-isolation, gray-virtues} to the
widespread adoption of loosely consistent distributed key-value
stores~\cite{bigtable, pnuts, dynamo}. Core to this shift was the 2000
introduction of Brewer's CAP Theorem, which stated that a highly
available system cannot provide ``strong'' consistency guarantees in
the presence of network partitions~\cite{brewer-slides}. As formally
proven~\cite{gilbert-cap}, the CAP Theorem pertains to a data
consistency model called linearizability, or the ability to read the
most recent write to a data item that is replicated across
servers~\cite{herlihy-art}. However, despite its narrow scope, the CAP
Theorem is often misconstrued as a broad result regarding the ability
to provide ACID database properties with high
availability~\cite{hat-hotos,brewer-slides, foundation-article}; this
misunderstanding has led to substantial confusion regarding replica
consistency, transactional isolation, and high availability. The
recent resurgence of transactional systems like Spanner~\cite{spanner}
suggests that programmers value transactional semantics, but,
unfortunately, existing transactional data stores do not provide
availability in the presence of partitions~\cite{orleans,foundation-article,
  hstore,spanner,eiger, walter,calvin}.

@@

To date, the relationship between ACID semantics and high
availability has not been well explored. We have a strong
understanding of weak isolation in the single-server context from
which it originated~\cite{adya, ansicritique, gray-isolation} and many
papers offer techniques for providing distributed
serializability~\cite{bernstein-book, spanner, daudjee-session,
  hstore, calvin} or snapshot
isolation~\cite{daudjee-snapshot, walter}. Additionally, the distributed computing and parallel
hardware literature contains many consistency models for single
operations on replicated objects~\cite{pnuts, herlihy-art, eiger, cac,
  sessionguarantees}. However, the literature lends few clues for
providing semantic guarantees for multiple operations operating on
multiple data items in a highly available distributed environment.

>>>>

The last decade has seen a shift in the design of popular large-scale
database systems, from the use of transactional
RDBMSs~\cite{bernstein-book, gray-isolation, gray-virtues} to the
widespread adoption of loosely consistent distributed key-value
stores~\cite{bigtable, pnuts, dynamo}. Core to this shift was the 2000
introduction of Brewer's CAP Theorem, which stated that a highly
available system cannot provide ``strong'' consistency guarantees in
the presence of network partitions~\cite{brewer-slides}. As formally
proven~\cite{gilbert-cap}, the CAP Theorem pertains to a data
consistency model called linearizability, or the ability to read the
most recent write to a data item that is replicated across
servers~\cite{herlihy-art}. However, despite its narrow scope, the CAP
Theorem is often misconstrued as a broad result regarding the ability
to provide ACID database properties with high
availability~\cite{hat-hotos,brewer-slides, foundation-article}; this
misunderstanding has led to substantial confusion regarding replica
consistency, transactional isolation, and high availability. The
recent resurgence of transactional systems suggests that programmers
value transactional semantics, but, unfortunately, existing
transactional data stores do not provide availability in the presence
of partitions~\cite{orleans,foundation-article, hstore,spanner,eiger,
  walter,calvin}.

@@

To date, the relationship between ACID semantics and high
availability has not been well explored. We have a strong
understanding of weak isolation in the single-server context from
which it originated~\cite{adya, ansicritique, gray-isolation} and many
papers offer techniques for providing distributed
serializability~\cite{bernstein-book, spanner, daudjee-session,
  hstore, calvin} or snapshot
isolation~\cite{kemme-thesis,walter}. Additionally, the distributed computing and parallel
hardware literature contains many consistency models for single
operations on replicated objects~\cite{pnuts, herlihy-art, eiger, cac,
  sessionguarantees}. However, the literature lends few clues for
providing semantic guarantees for multiple operations operating on
multiple data items in a highly available distributed environment.

### 2013-10-02 writeup/motivation.tex
More editing

Why does high availability matter? Peter Deutsch starts his classic
list of ``Fallacies of Distributed Computing'' with two concerns
fundamental to distributed database systems: ``\textit{1.)}  The
network is reliable. \textit{2.)} Latency is
zero''~\cite{fallacies-deutsch}; in a distributed setting, network
failures may prevent database servers from communicating, and, in the
absence of failures, communication is slowed by factors like physical
distance, network congestion, and routing. As we will see
(Section~\ref{sec:availability}), highly available system design
mitigates the effects of network \textit{partitions} and
\textit{latency}. In this section, we draw on a range of evidence that
indicates that partitions occur relatively frequently in real-world
deployments and best-case latencies between datacenters are
substantial, often on the order of several hundreds of milliseconds.

@@

Even with fault-free networks, distributed systems face the challenge
of network communication latency, Deutsch's second ``Fallacy.'' In
this section, we quantify round-trip latencies, which are often
large---hundreds of milliseconds in a geo-replicated, multi-datacenter
context.  Fundamentally, the speed at which two servers can
communicate is (according to modern physics) bounded by the speed of
light. In the best case, two servers on opposite sides of the
Earth---communicating via a hypothetical link along the planet's
diameter---require a minimum 85.1ms round-trip time (RTT; 133.7ms if
sent at surface level). As services are replicated to multiple,
geographically distinct sites, this cost of communication increases.

>>>>

Why does high availability matter? Peter Deutsch starts his classic
list of ``Fallacies of Distributed Computing'' with two concerns
fundamental to distributed database systems: ``\textit{1.)}  The
network is reliable. \textit{2.)} Latency is
zero''~\cite{fallacies-deutsch}; in a distributed setting, network
failures may prevent database servers from communicating, and, in the
absence of failures, communication is slowed by factors like physical
distance, network congestion, and routing. As we will see
(Section~\ref{sec:availability}), highly available system design
mitigates the effects of network \textit{partitions} and
\textit{latency}. In this section, we draw on a range of evidence that
indicates that partitions occur with frequency in real-world
deployments and latencies between datacenters are substantial, often
on the order of several hundreds of milliseconds.

@@

Even with fault-free networks, distributed systems face the challenge
of network communication latency, Deutsch's second ``Fallacy.'' In
this section, we quantify round-trip latencies, which are often
large---hundreds of milliseconds in a geo-replicated, multi-datacenter
context.  Fundamentally, the speed at which two servers can
communicate is (according to modern physics) bounded by the speed of
light. In the best case, two servers on opposite sides of the
Earth---communicating via a hypothetical link through the planet's
core---require a minimum 85.1ms round-trip time (RTT; 133.7ms if sent
at surface level). As services are replicated to multiple,
geographically distinct sites, this cost of communication increases.

### 2013-10-02 writeup/relatedwork.tex
More editing

There has been a recent resurgence of interest in distributed
multi-object semantics, both in academia~\cite{kraska-s3, gstore,
  eiger, walter,calvin, swift} and industry~\cite{orleans,spanner}. As
discussed in Section~\ref{sec:modernacid}, classic ACID databases
provide strong semantics but their lock-based and traditional
multi-versioned implementations are unavailable in the presence of
partitions~\cite{bernstein-book, gray-isolation}. Notably, Google's
Spanner provides strong one-copy serializable transactions. While
Spanner is highly specialized for Google's read-heavy workload, it
relies on two-phase commit and two-phase locking for read/write
transactions~\cite{spanner}. As we have discussed, the latency
penalties associated with this design choice are fundamental to
serializability. For users willing to tolerate these costs, Spanner,
or similar strongly consistent, unavailable
systems~\cite{kemme-classification}---including Calvin~\cite{calvin},
G-Store~\cite{gstore}, HBase, HStore~\cite{hstore},
Orleans~\cite{orleans}, Postgres-R~\cite{kemme-thesis},
Walter~\cite{walter}, and a range of snapshot isolation
techniques~\cite{daudjee-session, daudjee-snapshot}---are a reasonable
choice.

>>>>

There has been a recent resurgence of interest in distributed
multi-object semantics, both in academia~\cite{kraska-s3, gstore,
  eiger, walter,calvin, swift} and industry~\cite{orleans,spanner}. As
discussed in Section~\ref{sec:modernacid}, classic ACID databases
provide strong semantics but their lock-based and traditional
multi-versioned implementations are unavailable in the presence of
partitions~\cite{bernstein-book, gray-isolation}. Notably, Google's
Spanner provides strong one-copy serializable transactions. While
Spanner is highly specialized for Google's read-heavy workload, it
relies on two-phase commit and two-phase locking for read/write
transactions~\cite{spanner}. As we have discussed, the penalties
associated with this design are fundamental to serializability. For
users willing to tolerate unavailability and increased latency,
Spanner, or similar ``strongly consistent''
systems~\cite{kemme-classification}---including Calvin~\cite{calvin},
G-Store~\cite{gstore}, HBase, HStore~\cite{hstore},
Orleans~\cite{orleans}, Postgres-R~\cite{kemme-thesis},
Walter~\cite{walter}, and a range of snapshot isolation
techniques~\cite{daudjee-session}---reasonable
choices.
